---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

```{r, echo = TRUE}
rawdata <- read.csv("activity.csv")
data <- data.table(rawdata)
```

## What is mean total number of steps taken per day?

```{r ignoring missing values, echo = TRUE}
newdata <- data[complete.cases(data), ]
```
```{r calculating steps per day, echo = TRUE}
stepsperday <- as.numeric()
for (date in unique(newdata$date)) {
  stepsperday <- c(stepsperday, sum(newdata$steps[newdata$date == date]))
}
```
```{r histogram, echo = TRUE}
hist(stepsperday, main = "Steps per day")
```
```{r mean/median, echo = TURE}
mean(stepsperday)
median(stepsperday)
```

## What is the average daily activity pattern?

```{r echo = TRUE}
avsteps <- aggregate(steps ~ interval, newdata, mean)
plot(avsteps$interval, avsteps$steps, type="l", xlab = "interval", ylab = "steps")
avsteps[which.max(avsteps$steps), ]

```

## Imputing missing values

```{r echo = TRUE}
sum(is.na(data))
```
```{r filling missing values, echo = TRUE}
for (i in 1:nrow(data)) {
  if (is.na(data$steps[i])){
    int <- data$interval[i]
    row <- which(avsteps$interval == int)
    data$steps[i] <- avsteps$steps[row]
  }
}
```
```{r steps per day, echo=TRUE}
stepsperday2 <- as.numeric()
for (date in unique(newdata$date)) {
  stepsperday2 <- c(stepsperday2, sum(newdata$steps[newdata$date == date]))
}
hist(stepsperday2, main = "Steps per day")
mean(stepsperday2)
median(stepsperday2)

```
The mean is the same with or without missing values.
The median differs a very little.
By filling the missing values with the mean for intervals, the impact is practically unnoticeable.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
data$date <- as.character(data$date)
data$day <- weekdays(as.Date(data$date))
for (i in 1:nrow(data)) {
  if(data$day[i] == "sábado" || data$day[i] == "domingo"){
    data$day[i] <- "weekend"
  } else{
    data$day[i] <- "weekday"
  }
}
avsteps2 <- aggregate(steps ~ interval+day, data, mean)
library(ggplot2)
qplot(interval, steps, data=avsteps2, geom=c("line"), xlab="Interval", ylab="Steps", main="") + facet_wrap(~ day, ncol=1)
```